---
title: "01_DIABLO"
author: "Maria, zagor"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_caption: yes
    self_contained: yes
    fig_width: 12
    fig_height: 9
    toc: yes
    toc_depth: 5
    number_sections: yes
    theme: flatly
    highlight: tango
  pdf_document:
    toc: yes
    toc_depth: '5'
  word_document:
    toc: yes
    toc_depth: '5'
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(#dev = c('png', 'pdf', 'svg'),  # this embeds pdf and crates scrolable blocks
                      dev = c('png'), 
                      fig.align = 'center', 
                      fig.height = 9, 
                      fig.width = 12 ,
                      warning = FALSE, message = FALSE
                      )
# options(knitr.table.format = "html")

```



```{r}

gc()
rm(list = ls(all = TRUE))
gc()

set.seed(123456)

```





```{r}

# BiocManager::install("mixOmicsTeam/mixOmics@devel")
library(mixOmics)
set.seed(123456) 

# if (!requireNamespace("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")

# BiocManager::install("BiocParallel")
library(BiocParallel)



```

# Data

```{r}

fp = file.path('..', 'input')


fn = 'Enzymomics_only_24.txt'
E = read.delim(file = file.path(fp, fn), 
               header = TRUE, 
               sep = "\t", 
               quote = NULL,
               dec = ".", 
               fill = TRUE, 
               comment.char = '@')

fn = 'Metabolomics_only_24.txt'
M = read.delim(file = file.path(fp, fn), 
               header = TRUE, 
               sep = "\t", 
               quote = NULL,
               dec = ".", 
               fill = TRUE, 
               comment.char = '@')

fn = 'phosphoProteomics_filtered_sig_24.txt'
P = read.delim(file = file.path(fp, fn), 
               header = TRUE, 
               sep = "\t", 
               quote = NULL,
               dec = ".", 
               fill = TRUE, 
               comment.char = '@')

all(P$treatment == E$treatment)
all(P$treatment == M$treatment)

# cannot have that long names
colnames(M) = stringr::str_sub(colnames(M), -30) 



```


```{r}

data = list(Enz = as.matrix(E[, -1]),
            Met = as.matrix(M[, -1]),
            Pph = as.matrix(P[, -1]))



```

<https://mixomics.org/mixdiablo/diablo-tcga-case-study/>

```{r}


lapply(data, dim) # check their dimensions
Y = factor(E$treatment, levels = unique(E$treatment))
summary(Y)


```

# Pairwise PLS Comparisons

Circle Correlation Plots for pairwise PLS models 


Only displays the top 24 features for each dimension, subsetting by those with a correlation above 0.5

```{r}


list.keepX = c(24, 24) # select arbitrary values of features to keep
list.keepY = c(24, 24)

# generate three pairwise PLS models
pls1 <- spls(data[["Enz"]], data[["Met"]], 
             keepX = list.keepX, keepY = list.keepY) 
pls2 <- spls(data[["Enz"]], data[["Pph"]], 
             keepX = list.keepX, keepY = list.keepY)
pls3 <- spls(data[["Met"]], data[["Pph"]], 
             keepX = list.keepX, keepY = list.keepY)

# plot features of first PLS
plotVar(pls1, cutoff = 0.5, title = "(a) enzymomics vs metabolomics", 
        legend = c("enzymomics", "metabolomics"), 
        var.names = TRUE, style = 'graphics', 
        pch = c(16, 17), cex = c(0.5,0.5), 
        col = c('darkorchid', 'lightgreen'))
plotVar(pls1, cutoff = 0.5, title = "(a) enzymomics vs metabolomics", 
        legend = c("enzymomics", "metabolomics"), 
        var.names = FALSE, style = 'graphics', 
        pch = c(16, 17), cex = c(2,2), 
        col = c('darkorchid', 'lightgreen'))

# plot features of second PLS
plotVar(pls2, cutoff = 0.5, title = "(b) enzymomics vs phosphoproteomics", 
        legend = c("enzymomics", "phosphoproteomics"), 
        var.names = TRUE, style = 'graphics', 
        pch = c(16, 17), cex = c(0.5,0.5), 
        col = c('darkorchid', 'blue'))
plotVar(pls2, cutoff = 0.5, title = "(b) enzymomics vs phosphoproteomics", 
        legend = c("enzymomics", "phosphoproteomics"), 
        var.names = FALSE, style = 'graphics', 
        pch = c(16, 17), cex = c(2,2), 
        col = c('darkorchid', 'blue'))

# plot features of third PLS
plotVar(pls3, cutoff = 0.5, title = "(c) metabolomics vs phosphoproteomics", 
        legend = c("metabolomics", "phosphoproteomics"), 
        var.names = TRUE, style = 'graphics', 
        pch = c(16, 17), cex = c(0.5,0.5), 
        col = c('lightgreen', 'blue'))
plotVar(pls3, cutoff = 0.5, title = "(c) metabolomics vs phosphoproteomics", 
        legend = c("metabolomics", "phosphoproteomics"), 
        var.names = FALSE, style = 'graphics', 
        pch = c(16, 17), cex = c(2,2), 
        col = c('lightgreen', 'blue'))


```

## correlation

```{r}

cor(pls1$variates$X, pls1$variates$Y) # calculate correlation of Enz and Met
cor(pls2$variates$X, pls2$variates$Y) # calculate correlation of Enz and Pph
cor(pls3$variates$X, pls3$variates$Y) # calculate correlation of Met and Pph

```


# Design



```{r}

design = matrix(0.5, ncol = length(data), nrow = length(data), # for square matrix filled with 0.1s
                dimnames = list(names(data), names(data)))
diag(design) = 0 # set diagonal to 0s

design

```

#  the initial DIABLO 


```{r}

basic.diablo.model = block.splsda(X = data, Y = Y, ncomp = 5, design = design) # form basic DIABLO model

```


# Tuning the number of components

Choosing the number of components in block.plsda using perf() with 10 × 10-fold CV function 

```{r}
## ---- fig.cap = "FIGURE 2: Choosing the number of components in `block.plsda` using `perf()` with 10 × 10-fold CV function in the `breast.TCGA` study. Classification error rates (overall and balanced, see Section 7.3) are represented on the y-axis with respect to the number of components on the x-axis for each prediction distance presented in PLS-DA"----

perf.diablo = perf(basic.diablo.model, validation = 'Mfold', 
                   folds = 10, nrepeat = 10) # run component number tuning with repeated CV

plot(perf.diablo) # plot output of tuning

```


##  set the optimal ncomp value


```{r}


ncomp = perf.diablo$choice.ncomp$WeightedVote["Overall.BER", "centroids.dist"] # set the optimal ncomp value
perf.diablo$choice.ncomp$WeightedVote # show the optimal choice for ncomp for each dist metric

```

# Tuning the number of features

We choose the optimal number of variables to select in each data set using the tune.block.splsda() function, for a grid of keepX values for each type of omics. Note that the function has been set to favour a relatively small signature while allowing us to obtain a sufficient number of variables for downstream validation and/or interpretation.


## setup cluster - use SnowParam() on Widnows

<https://rdrr.io/bioc/BiocParallel/man/BiocParallelParam-class.html>


bplapply erro solution <https://support.bioconductor.org/p/133353/#133356>

```{r}

options(SnowParam=SnowParam(workers=4))
BPPARAM  = bpparam()




```


Crate tune.MIR with tune.block.splsda

and save RData

then comment these lines

and just load RData

```{r}

# set grid of values for each component to test


 # test.keepX = list (Met = c(5:9, seq(10, 15, 2), seq(20,24,4)),
 #                    Enz = c(5:9, seq(10, 15, 2), seq(20,24,4)),
 #                    Pph = c(5:9, seq(10, 15, 2), seq(20,24,4)))

x <- list()
for (i in 1:length(data)){
x[[i]] <- c( seq(5,min(24, ncol(data[[i]])) ,2))
}
names(x) <- names(data)
test.keepX  <-  x
test.keepX


# You have provided a sequence of keepX of length: 10 for block Met and 10 for block Enz and 10 for block Pph.
# This results in 1000 models being fitted for each component and each nrepeat, this may take some time to run, be patient!
# Because of a too high number of 'folds' required, 2 folds were randomly assigned no data: the number of 'folds' is reduced to 8
 
 t1 = proc.time()
 
 # run the feature selection tuning
 
 # tune.MIR = tune.block.splsda(X = data, Y = Y, ncomp = ncomp,
 #                               test.keepX = test.keepX, design = design,
 #                              validation = 'Mfold', folds = 8, nrepeat = 1,
 #                              BPPARAM  = BPPARAM , dist = "centroids.dist")
#  t2 = proc.time()
#  running_time = t2 - t1; running_time
# 
# list.keepX = tune.MIR$choice.keepX
# list.keepX
# 
# 
fp = file.path('..', 'output')
fn = 'result-MIR-diablo_design0.1.RData'
# save(tune.MIR,list.keepX, file = file.path(fp, fn))
 
load(file.path(fp, fn))

```


The number of features to select on each component is returned in


```{r}



list.keepX = tune.MIR$choice.keepX # set the optimal values of features to retain
list.keepX


```

# Final model

```{r}


# set the optimised DIABLO model
final.diablo.model = block.splsda(X = data, Y = Y, ncomp = ncomp
                          , keepX = list.keepX
                          , design = design)

## Design matrix has changed to include Y; each block will be
##             linked to Y.



```



```{r}

# the features selected from components
for (comp in 1:ncomp){
cat("\nComponent ", comp,":\n")
for(i in 1:length(data)){
cat(names(data)[i],"\n")
print(selectVar(final.diablo.model, comp = comp)[[i]]$name)
}
}



```


# Sample plots


```{r}


for(comp in 1:ncomp){
plotDiablo(final.diablo.model, ncomp = comp)
title(paste("Component",comp), adj=0.1, line=-1, outer=TRUE)
}
```


```{r}


plind <- plotIndiv(final.diablo.model, ind.names = FALSE, legend = TRUE,
          title = 'DIABLO Sample Plots'
          , ellipse = TRUE
          )



```


```{r}

plotArrow(final.diablo.model, ind.names = FALSE, legend = TRUE,
          title = paste(groups,collapse=", ")

          )



```


# Variable plots

```{r}

if(length(data)==3) pick <- 1:3 else pick <- c(4,1:3)
cols <- c('orange1', 'brown1', 'lightgreen',"lightblue")[pick]
pchs <- c(16, 17, 15, 18)[pick]
plotVar(final.diablo.model, var.names = FALSE,
        style = 'graphics', legend = TRUE
        , pch = pchs, cex = rep(2,length(data))
        , col = cols
)

plotVar(final.diablo.model, var.names = TRUE,
        style = 'graphics', legend = TRUE
        , pch = pchs, cex = rep(0.5,length(data))
        , col = cols
)

```


```{r}

cutoff = 0.75

circosPlot(final.diablo.model, cutoff = cutoff, line = TRUE,
           color.blocks= cols,
           color.cor = c(3,2), size.labels = 1
           , xpd=TRUE)




```

# Relevance networks



```{r}

# png(paste0(1,".png"), res = 600, width = 4000, height = 4000)
# nw <- network(final.diablo.model
#         , blocks = blocks[,j]
#         , color.node = cols[blocks[,j]]
#         , cutoff = cutoff
#         , shape.node = "rectangle"
#         , save = "png"
#        , name.save = nfn
#         )
# 
# png(paste0(2,".png"), res = 600, width = 4000, height = 4000)
# nw <- network(final.diablo.model
#         , blocks = blocks[,j]
#         , color.node = cols[blocks[,j]]
#         , cutoff = cutoff
#         , shape.node = "rectangle"
#         , save = "png"
#        , name.save = nfn
#         )

# nw <- my.network(x
#         , blocks = 1:ndata
#         , color.node = cols
#         , cutoff = cutoff
#         , shape.node = "rectangle"
#         , layout = layout.fun
#         , save = "png"
#         , name.save = 'test'
#         )

```


# more plots

issue <https://github.com/mixOmicsTeam/mixOmics/issues/45>

```{r}


for(i in 1:ncomp)
plotLoadings(final.diablo.model, comp = i, contrib = 'max', method = 'median')


# plotLoadings encountered margin errors. Ensure feature names are not too long (see 'name.var' argument) and the 'Plots' pane is cleared and enlargened.


traceback()


```

```{r}

cimfn <- "cim.png"
png(cimfn, res = 600, width = 8000, height = 4000)
cimDiablo(final.diablo.model, size.legend=0.7)
dev.off()



```

# Heatmap

```{r}
#heatmaps
cimDiablo(final.diablo.model)
```


# Multipartite network

```{r network}

cutoff = 0.75

#cutoff for correlations
network(final.diablo.model, blocks = c(1,2,3),
        color.node = c('darkorchid', 'brown1', 'lightgreen'), 
        cutoff = cutoff, 
        size.node = 0.1,
        color.edge = c('darkblue', 'blue', 'red', 'darkred'))
```



An AUC plot per block can also be obtained using the function auroc(). The interpretation of this output may not be particularly insightful in relation to the performance evaluation of our methods, but can complement the statistical analysis..



```{r AUC}

for(i in 1:length(data))  {
  auc.splsda = auroc(final.diablo.model, roc.block = names(data[i]),
                     roc.comp = 2, print = FALSE)
}



```


# Condition-specific networks

```{r}

library(igraph)

```


```{r}

# taken from  005-Functions.Rnw/005-Functions.Rmd

extractEdges2 <- function(x, cnames=NULL){

    edges <- NULL
    ind <- grep("^M_",names(x))
    for (i in ind){
    e <- x[[i]]
    from <- rep(rownames(e), ncol(e))
    to <- rep(colnames(e), each=nrow(e))
    group <- names(x)[i]
    w <- e[from, to]
    group1 <- gsub("^M_(.*)_(.*)$","\\1",group)
    gr1 <- substr(group1,1,2)
    group2 <- gsub("^M_(.*)_(.*)$","\\2",group)
    gr2 <- substr(group2,1,2)
        lbl <-     paste(paste(gr1,from,sep="."),paste(gr2,to,sep="."),sep="_")
        es <- data.frame(edge=lbl, group1=group1
        , from=from
        , group2=group2
        , to=to
        , w=as.vector(e))
#        print(head(es))
#        print(head(e))
    rownames(es) <- paste(paste(gr1,from,sep="."),paste(gr2,to,sep="."),sep="_")
    edges <- rbind(edges,es)
    }
    if(!is.null(cnames)) colnames(edges) <- cnames
    edges

  }    
    
```



```{r}

CCDATA = data
names(CCDATA)

CCDATA$status = Y



```


## NM_24 & NM_inf_24

```{r}

(whichOne = levels(Y)[1:2])


filter =  CCDATA$status  %in% whichOne
XX1 <- lapply(CCDATA, function(x) if(is.null(dim(x))) x[filter] else x[filter,])
table(XX1$status)

res1 <- block.splsda(X = XX1[-length(XX1)]
    , Y = as.factor(XX1[[length(XX1)]])
    , ncomp = ncomp
    , keepX = list.keepX
    , design = design
    )

cutoff = 0.0
x = res1
layout.fun <- NULL
label = whichOne

ndata = length(data)
lbl =  gsub(", ","-",label)

nw1 = network(x
        , blocks = 1:ndata
        , color.node = c('darkorchid', 'brown1', 'lightgreen')
        , cutoff = 0.7# cutoff
        , shape.node = "rectangle"
        , layout = layout_nicely
        , save = "png"
        , size.node = 0.1
        , name.save = paste(lbl, collapse = ' vs ')
        )


e1 = extractEdges2(nw1)
colnames(e1)[ncol(e1)] <- paste(whichOne, collapse=".")
head(e1)



```


## AM_24 & AM_inf_24

```{r}

(whichOne = levels(Y)[3:4])

filter =  CCDATA$status  %in% whichOne
XX2 <- lapply(CCDATA, function(x) if(is.null(dim(x))) x[filter] else x[filter,])
table(XX2$status)

res2 <- block.splsda(X = XX2[-length(XX2)]
    , Y = as.factor(XX2[[length(XX2)]])
    , ncomp = ncomp
    , keepX = list.keepX
    , design = design
    )

cutoff = 0.0
x = res2
layout.fun <- NULL
label = whichOne

ndata = length(data)
lbl =  gsub(", ","-",label)

nw2 = network(x
        , blocks = 1:ndata
        , color.node = c('darkorchid', 'brown1', 'lightgreen')
        , cutoff = 0.7# cutoff
        , shape.node = "rectangle"
        , layout = layout_nicely
        , save = "png"
        , size.node = 0.1
        , name.save = paste(lbl, collapse = ' vs ')
        )

e2 = extractEdges2(nw2)
colnames(e2)[ncol(e2)] <- paste(whichOne, collapse=".")
head(e2)



```

## NM_24 & AM_24

```{r}

(whichOne = levels(Y)[c(1, 3)])

filter =  CCDATA$status  %in% whichOne
XX3 <- lapply(CCDATA, function(x) if(is.null(dim(x))) x[filter] else x[filter,])
table(XX3$status)

res3 <- block.splsda(X = XX3[-length(XX3)]
    , Y = as.factor(XX3[[length(XX3)]])
    , ncomp = ncomp
    , keepX = list.keepX
    , design = design
    )

cutoff = 0.0
x = res3
layout.fun <- NULL
label = whichOne

ndata = length(data)
lbl =  gsub(", ","-",label)

nw3 = network(x
        , blocks = 1:ndata
        , color.node = c('darkorchid', 'brown1', 'lightgreen')
        , cutoff = 0.7# cutoff
        , shape.node = "rectangle"
        , layout = layout_nicely
        , save = "png"
        , size.node = 0.1
        , name.save = paste(lbl, collapse = ' vs ')
        )

e3 = extractEdges2(nw3)
colnames(e3)[ncol(e3)] <- paste(whichOne, collapse=".")
head(e3)



```



## NM_24 & AM_24_inf

```{r}

(whichOne = levels(Y)[c(1, 4)])

filter =  CCDATA$status  %in% whichOne
XX4 <- lapply(CCDATA, function(x) if(is.null(dim(x))) x[filter] else x[filter,])
table(XX4$status)

res4 <- block.splsda(X = XX4[-length(XX4)]
    , Y = as.factor(XX4[[length(XX4)]])
    , ncomp = ncomp
    , keepX = list.keepX
    , design = design
    )

cutoff = 0.0
x = res4
names(x)
range(x$weights)
layout.fun <- NULL
label = whichOne

ndata = length(data)
lbl =  gsub(", ","-",label)

nw4 = network(x
        , blocks = 1:ndata
        , color.node = c('darkorchid', 'brown1', 'lightgreen')
        , cutoff = 0.7# cutoff
        , shape.node = "rectangle"
        , layout.fun = layout_nicely
        , save = "png"
        , size.node = 0.1
        , name.save = paste(lbl, collapse = ' vs ')
        )

e4 = extractEdges2(nw4)
colnames(e4)[ncol(e4)] <- paste(whichOne, collapse=".")
head(e4)



```


## NM_24_inf & AM_24_inf

```{r}

(whichOne = levels(Y)[c(2, 4)])

filter =  CCDATA$status  %in% whichOne
XX5 <- lapply(CCDATA, function(x) if(is.null(dim(x))) x[filter] else x[filter,])
table(XX5$status)

res5 <- block.splsda(X = XX5[-length(XX5)]
    , Y = as.factor(XX5[[length(XX5)]])
    , ncomp = ncomp
    , keepX = list.keepX
    , design = design
    )

cutoff = 0.0
x = res5
names(x)
range(x$weights)
layout.fun <- NULL
label = whichOne

ndata = length(data)
lbl =  gsub(", ","-",label)

# Relevance Network for (r)CCA and (s)PLS regression
nw5 = network(x
        , blocks = 1:ndata
        , color.node = c('darkorchid', 'brown1', 'lightgreen')
        , cutoff = 0.7# cutoff
        , shape.node = "rectangle"
        # , layout.fun = layout_on_grid 
        , layout = layout.fun
        , save = "png"
        , size.node = 0.1
        , name.save = paste(lbl, collapse = ' vs ')
        , plot.graph = FALSE
        )

e5 = extractEdges2(nw5)
colnames(e5)[ncol(e6)] <- paste(whichOne, collapse=".")
head(e5)



```




# sessionInfo

```{r}

sessionInfo()

```


# See also comments at

 ```GitHub/UJI-Omics-phd/_I_UJI-Omics-phd/_S_multiOmicsIntegra/_A_10_mixOmics-R/scripts/03_Step5.html```



