---
title: "02_DIABLO_5omics_levels"
author: "Maria, zagor"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_caption: yes
    self_contained: yes
    fig_width: 12
    fig_height: 9
    toc: yes
    toc_depth: 2
    number_sections: yes
    theme: flatly
    highlight: tango
  pdf_document:
    toc: yes
    toc_depth: '2'
  word_document:
    toc: yes
    toc_depth: '2'
editor_options:
  chunk_output_type: console
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(#dev = c('pdf', 'png'),  # this embeds pdf and crates scrolable blocks
                      dev = c('png', 'svg', 'pdf'), 
                      fig.align = 'center', 
                      fig.height = 9, 
                      fig.width = 12 ,
                      warning = FALSE, message = FALSE
                      )
# options(knitr.table.format = "html")

```

```{r}

gc()
rm(list = ls(all = TRUE))
gc()

set.seed(123456)

library(igraph)

```


fix layout error

```{r, echo=FALSE}

bin.color = function(mat, cutoff, breaks, col, symkey) 
{
    if (isTRUE(symkey)) {
        max.mat = max(abs(mat))
        min.mat = -max.mat
    }
    else {
        max.mat = max(mat)
        min.mat = min(mat)
    }
	
	    if (missing(breaks) || is.null(breaks)) {
        if (class(col) == "function") breaks = 32
            else breaks = length(col) 
    }

    if (length(breaks) == 1) {
		if (isTRUE(symkey)) {
				if ((breaks/2) - trunc(breaks/2) != 0) 
					stop("'breaks' must be a even number if 'symkey = TRUE'", call. = FALSE)
			
			if (cutoff == 0) breaks = c(seq(min.mat, max.mat, length = breaks + 1))
			else {			
				nb = breaks/2
				breaks = c(seq(min.mat, -cutoff, length = nb + 1), 0, 
						   seq(cutoff, max.mat, length = nb + 1))					   
				id = which(breaks == 0)
				breaks = breaks[-c(id - 1, id + 1)]			
			}
		}
		else { 
			breaks = breaks + 1
			
			if ((min.mat < -cutoff) & (max.mat < cutoff))
				breaks = seq(min.mat, -cutoff, length = breaks)
				
			if ((min.mat > -cutoff) & (max.mat > cutoff))
				breaks = seq(cutoff, max.mat, length = breaks)
				
			if ((min.mat < -cutoff) & (max.mat > cutoff)) {
				if (cutoff == 0) breaks = c(seq(min.mat, max.mat, length = breaks))
				else {
					long = max.mat - min.mat - 2*cutoff
					bin = long/breaks
					breaks = seq(cutoff, -min.mat, by = bin)
					o = order(breaks, decreasing = TRUE)
					breaks = c(-breaks[o], 0, seq(cutoff, max.mat, by = bin))
					id = which(breaks == 0)				
					breaks = breaks[-c(id - 1, id + 1)]
				}
			}
		}
    }

    ncol = length(breaks) - 1

    if (class(col) == "function") 
        col = col(ncol)
		
	if (length(breaks) != length(col) + 1) 
        stop("must have one more break than colour", call. = FALSE)	

    min.breaks = min(breaks)
    max.breaks = max(breaks)

    mat[mat < min.breaks] = min.breaks
    mat[mat > max.breaks] = max.breaks
    
    bin = .bincode(as.double(mat), as.double(breaks), TRUE, TRUE)

    return(invisible(list(bin = bin, col = col, breaks = breaks, lim = c(min.mat, max.mat))))		
}


network <- function (mat, comp = NULL, blocks = c(1, 2), cutoff = 0, row.names = TRUE, 
    col.names = TRUE, block.var.names = TRUE, graph.scale = 0.5, 
    size.node = 0.5, color.node = NULL, shape.node = NULL, alpha.node = 0.85, 
    cex.node.name = NULL, color.edge = color.GreenRed(100), lty.edge = "solid", 
    lwd.edge = 1, show.edge.labels = FALSE, cex.edge.label = 1, 
    show.color.key = TRUE, symkey = TRUE, keysize = c(1, 1), 
    keysize.label = 1, breaks, interactive = FALSE, layout.fun = NULL, 
    save = NULL, name.save = NULL, plot.graph = TRUE) 
{
    arg.call = match.call()
    user.arg = names(arg.call)[-1]
    err = tryCatch(mget(names(formals()), sys.frame(sys.nframe())), 
        error = function(e) e)
    if ("simpleError" %in% class(err)) 
        stop(err[[1]], ".", call. = FALSE)
    function.arg = names(mget(names(formals()), sys.frame(sys.nframe())))
    not.arg = !(user.arg %in% function.arg)
    if (any(not.arg)) {
        unused.arg = user.arg[not.arg]
        not.arg = which(not.arg) + 1
        output = rep("", length(not.arg))
        for (i in 1:length(not.arg)) {
            output[i] = paste0(unused.arg[i], " = ", arg.call[[not.arg[i]]])
        }
        output = paste0("(", paste(output, collapse = ", "), 
            ").")
        msg = "unused argument "
        if (length(not.arg) > 1) 
            msg = "unused arguments "
        stop(msg, output, call. = FALSE)
    }
    if (!is.null(save)) {
        if (!save %in% c("jpeg", "tiff", "png", "pdf")) 
            stop("'save' must be one of 'jpeg', 'png', 'tiff' or 'pdf'.", 
                call. = FALSE)
    }
    if (!is.null(name.save)) {
        if (!is.character(name.save) || length(name.save) > 1) 
            stop("'name.save' must be a character.", call. = FALSE)
    }
    else {
        if (!is.null(save)) 
            name.save = paste0("network_", gsub(".", "_", deparse(substitute(mat)), 
                fixed = TRUE))
    }
    if (!is.null(save)) {
        while (dev.cur() > 1) dev.off()
        if (save == "jpeg") 
            jpeg(filename = paste0(name.save, ".jpeg"), res = 600, 
                width = 4000, height = 4000)
        if (save == "png") 
            jpeg(filename = paste0(name.save, ".png"), res = 600, 
                width = 4000, height = 4000)
        if (save == "tiff") 
            tiff(filename = paste0(name.save, ".tiff"), res = 600, 
                width = 4000, height = 4000)
        if (save == "pdf") 
            pdf(file = paste0(name.save, ".pdf"))
    }
    if (!plot.graph & interactive) {
        stop("plot.graph cannot be FALSE if interactive = TRUE", 
            call. = FALSE)
    }
    class.object = class(mat)
    object.pls = c("mixo_pls", "mixo_spls", "mixo_mlspls")
    object.rcc = "rcc"
    object.blocks = c("sgcca", "rgcca")
    if (!any(class.object %in% c(object.pls, object.rcc, object.blocks, 
        "matrix"))) 
        stop(" 'network' is only implemented for the following objects: matrix, pls, plsda, spls, splsda, rcc, sgcca, rgcca, sgccda", 
            call. = FALSE)
    if (any(class.object %in% c(object.rcc, object.pls))) {
        p = ncol(mat$X)
        if (any(class.object == "DA")) 
            mat$Y = mat$ind.mat
        q = ncol(mat$Y)
        n = nrow(mat$X)
        ncomp = mat$ncomp
        if (is.null(comp)) 
            comp = 1:mat$ncomp
        if (length(comp) == 1) {
            if (comp > ncomp) {
                stop("the elements of 'comp' must be smaller than or equal to ", 
                  ncomp, ".", call. = FALSE)
            }
            else if (!is.numeric(comp) || comp <= 0) {
                stop("invalid value for 'comp'.", call. = FALSE)
            }
        }
        if (length(comp) > 1) {
            if (length(comp) > ncomp) 
                stop("the length of 'comp' must be smaller than or equal to ", 
                  ncomp, ".", call. = FALSE)
            if (!is.numeric(comp) || any(comp < 1)) 
                stop("invalid vector for 'comp'.", call. = FALSE)
            if (any(comp > ncomp)) 
                stop("the elements of 'comp' must be <= ", ncomp, 
                  ".", call. = FALSE)
        }
        comp = round(comp)
        row.names.plot = TRUE
        if (is.logical(row.names)) {
            if (!isTRUE(row.names)) {
                row.names.plot = FALSE
            }
            row.names = mat$names$colnames$X
        }
        else {
            row.names = as.vector(row.names)
            if (length(unique(row.names)) != p) 
                stop("'row.names' must be a character vector of ", 
                  p, " unique entries.", call. = FALSE)
        }
        if (row.names.plot == TRUE) {
            row.names.plot = row.names
        }
        else {
            row.names.plot = rep("", p)
        }
        col.names.plot = TRUE
        if (is.logical(col.names)) {
            if (!isTRUE(col.names)) {
                col.names.plot = FALSE
            }
            col.names = mat$names$colnames$Y
        }
        else {
            col.names = as.vector(col.names)
            if (length(col.names) != q) 
                stop("'col.names' must be a character vector of ", 
                  q, " unique entries.", call. = FALSE)
        }
        if (col.names.plot == TRUE) {
            col.names.plot = col.names
        }
        else {
            col.names.plot = rep("", q)
        }
        if (any(class.object %in% object.rcc)) {
            bisect = mat$variates$X[, comp] + mat$variates$Y[, 
                comp]
            cord.X = cor(mat$X, bisect, use = "pairwise")
            cord.Y = cor(mat$Y, bisect, use = "pairwise")
            mat = cord.X %*% t(cord.Y)
        }
        else if (any(class.object %in% object.pls)) {
            if (all(class(mat) %in% "mixo_pls")) {
                keep.X = rep(TRUE, p)
                keep.Y = rep(TRUE, q)
            }
            else {
                keep.X = apply(abs(mat$loadings$X[, comp, drop = FALSE]), 
                  1, sum) > 0
                keep.Y = apply(abs(mat$loadings$Y[, comp, drop = FALSE]), 
                  1, sum) > 0
                row.names = row.names[keep.X]
                col.names = col.names[keep.Y]
            }
            if (mat$mode == "canonical") {
                cord.X = cor(mat$X[, keep.X], mat$variates$X[, 
                  comp], use = "pairwise")
                cord.Y = cor(mat$Y[, keep.Y], mat$variates$Y[, 
                  comp], use = "pairwise")
            }
            else {
                cord.X = cor(mat$X[, keep.X], mat$variates$X[, 
                  comp], use = "pairwise")
                cord.Y = cor(mat$Y[, keep.Y], mat$variates$X[, 
                  comp], use = "pairwise")
            }
            mat = cord.X %*% t(cord.Y)
        }
    }
    else if (any(class.object %in% object.blocks)) {
        mat$X <- mapply(mat$X, names(mat$X), FUN = function(x, 
            y) {
            colnames(x) <- paste0(colnames(x), "_", y)
            x
        }, SIMPLIFY = FALSE)
        if (any(class.object == "DA")) {
            mat$names$blocks = mat$names$blocks[-mat$indY]
            mat$names$colnames = mat$names$colnames[-mat$indY]
            mat$ncomp = mat$ncomp[-mat$indY]
        }
        if (is.null(blocks)) {
            if (any(mat$ncomp > 1)) {
                blocks = mat$names$blocks[which(mat$ncomp > 1)]
            }
            else {
                stop(("The number of components for each block is 1. The number of components must be superior or equal to 2."), 
                  call. = FALSE)
            }
        }
        else if (is.numeric(blocks) & min(blocks) > 0 & max(blocks) <= 
            length(mat$names$blocks)) {
            blocks = mat$names$blocks[blocks]
        }
        else if (is.character(blocks)) {
            if (!all(blocks %in% mat$names$blocks)) 
                stop("One element of 'blocks' does not match with the names of the blocks")
        }
        else {
            stop("Incorrect value for 'blocks", call. = FALSE)
        }
        if (is.null(comp)) {
            comp = vector("list", length(blocks))
            names(comp) = blocks
            for (i in blocks) comp[[i]] = 1:mat$ncomp[i]
        }
        if (is.list(comp)) {
            if (length(comp) != length(blocks)) 
                stop("'comp' must be either NULL a list of length ", 
                  length(blocks), ".", call. = FALSE)
            if (!all(blocks %in% names(comp))) 
                stop("names of 'comp' must be from {", paste(blocks, 
                  collapse = ", "), "}.", call. = FALSE)
            for (i in blocks) {
                if (any(!is.finite(comp[[i]]))) 
                  stop("invalid value for 'comp' of the block '", 
                    i, "'.", call. = FALSE)
                if (any(comp[[i]] > mat$ncomp[i])) 
                  stop("the elements of 'comp' for block '", 
                    i, "' must be <= ", mat$ncomp[i], ".", call. = FALSE)
                if (any(comp[[i]] < 1)) 
                  stop("invalid value for 'comp' of the block '", 
                    i, "'.", call. = FALSE)
            }
        }
        else {
            stop("'comp' must be either NULL or a list of length ", 
                length(blocks), ".", call. = FALSE)
        }
        num.var = unlist(lapply(mat$X[blocks], ncol))
        if (is.logical(block.var.names)) {
            if (length(block.var.names) == 1) 
                block.var.names = rep(block.var.names, length(blocks))
            if (length(block.var.names) != length(blocks)) 
                stop("'block.var.names' must be a logical vector of length 1 or ", 
                  length(blocks), ", or a list of length ", length(blocks), 
                  ".", call. = FALSE)
            vec = (which(block.var.names == FALSE))
            block.var.names = mat$names$colnames
            for (i in 1:length(blocks)) {
                if (i %in% vec) 
                  block.var.names[[blocks[i]]] = rep(" ", length(mat$names$colnames[[blocks[i]]]))
            }
        }
        else {
            if (is.list(block.var.names)) {
                if (length(block.var.names) != length(blocks)) {
                  stop("'block.var.names' must be a logical vector or a list of length ", 
                    length(blocks), ".", call. = FALSE)
                }
                else {
                  if (!all(unlist(lapply(block.var.names, is.vector)))) 
                    stop("each component of 'block.var.names' must be a vector.", 
                      call. = FALSE)
                  block.var.names.length = unlist(lapply(block.var.names, 
                    length))
                  if (any(block.var.names.length != num.var)) 
                    stop("components of 'block.var.names' must be vectors of length ", 
                      paste(num.var, collapse = ", "), ".", call. = FALSE)
                }
            }
            else {
                stop("'block.var.names' must be either a logical value or a list of length ", 
                  length(blocks), ".", call. = FALSE)
            }
        }
        coord = M_block = list()
        j = 1
        if (is(mat, "sgcca")) {
            for (k in blocks) {
                if (length(comp[[k]]) > 1) {
                  keep = (apply(abs(mat$loadings[[k]][, comp[[k]]]), 
                    1, sum) > 0)
                }
                else {
                  keep = abs(mat$loadings[[k]][, comp[[k]]]) > 
                    0
                }
                coord[[j]] = cor(mat$X[[k]][, keep], mat$variates[[k]][, 
                  comp[[k]]], use = "pairwise")
                j = j + 1
            }
        }
        else {
            for (k in blocks) {
                coord[[j]] = cor(mat$X[[k]], mat$variates[[k]][, 
                  comp[[k]]], use = "pairwise")
                j = j + 1
            }
        }
        node.X = node.Y = w = NULL
        l = 1
        for (j in 1:(length(blocks) - 1)) {
            for (k in (j + 1):length(blocks)) {
                if (!any(comp[[blocks[j]]] %in% comp[[blocks[k]]])) 
                  stop("comp of block ", blocks[j], " is ", comp[[blocks[j]]], 
                    " but comp of block ", blocks[k], " is ", 
                    comp[[blocks[k]]], call. = FALSE)
                int.comp = intersect(comp[[blocks[j]]], comp[[blocks[k]]])
                object = coord[[j]][, comp[[blocks[j]]] %in% 
                  int.comp] %*% t(coord[[k]][, comp[[blocks[k]]] %in% 
                  int.comp])
                M_block[[l]] = object
                l = l + 1
                X = rownames(coord[[j]])
                Y = rownames(coord[[k]])
                rep.X = rep(X, each = length(Y))
                rep.Y = rep(Y, length(X))
                node.X = c(node.X, rep.X)
                node.Y = c(node.Y, rep.Y)
                w = c(w, as.vector(t(object)))
            }
        }
    }
    else {
        if (!is.matrix(mat)) 
            stop("'mat' must be a numeric matrix.", call. = FALSE)
        if (length(dim(mat)) != 2) 
            stop("'mat' must be a numeric matrix.")
        if (!is.numeric(mat)) 
            stop("'mat' must be a numeric matrix.")
        p = nrow(mat)
        q = ncol(mat)
        row.names.plot = TRUE
        if (is.logical(row.names)) {
            if (!isTRUE(row.names)) {
                row.names.plot = FALSE
            }
            row.names = rownames(mat)
        }
        else {
            row.names = as.vector(row.names)
            if (length(row.names) != p) 
                stop("'row.names' must be a character vector of ", 
                  p, " unique entries.", call. = FALSE)
        }
        if (row.names.plot == TRUE) {
            row.names.plot = row.names
        }
        else {
            row.names.plot = rep("", p)
        }
        col.names.plot = TRUE
        if (is.logical(col.names)) {
            if (!isTRUE(col.names)) {
                col.names.plot = FALSE
            }
            col.names = colnames(mat)
        }
        else {
            col.names = as.vector(col.names)
            if (length(col.names) != q) 
                stop("'col.names' must be a character vector of ", 
                  q, " unique entries.", call. = FALSE)
        }
        if (col.names.plot == TRUE) {
            col.names.plot = col.names
        }
        else {
            col.names.plot = rep("", q)
        }
    }
    if (!is.null(size.node)) {
        if (!is.finite(size.node) || size.node < 0 || size.node > 
            1) {
            stop("'size.node' must be a numerical value between 0 - 1.", 
                call. = FALSE)
        }
        if (size.node == 0) {
            shape.node <- "none"
            size.node <- 1
        }
    }
    if (!is.finite(graph.scale) || graph.scale < 0 || graph.scale > 
        1) {
        stop("'graph.scale' must be a numerical value between 0 - 1.", 
            call. = FALSE)
    }
    if (any(class.object %in% object.blocks)) {
        if (is.null(color.node)) {
            color.blocks = brewer.pal(n = 12, name = "Paired")
            if (length(blocks) > 6) {
                color.blocks <- colorRampPalette(color.blocks)(2 * 
                  length(object$X))
            }
            color.node = color.blocks[seq(from = 1, to = 2 * 
                length(blocks), by = 2)]
            color.node = adjustcolor(color.node, alpha.f = alpha.node)
            names(color.node) = blocks
        }
        else {
            if (!is.vector(color.node) || length(color.node) != 
                length(blocks)) 
                stop("'color.node' must be a vector of length ", 
                  length(blocks), ".", call. = FALSE)
        }
    }
    else {
        if (is.null(color.node)) 
            color.node = brewer.pal(n = 12, name = "Paired")[c(1, 
                3)]
        color.node = adjustcolor(color.node, alpha.f = alpha.node)
        if (!is.list(color.node)) {
            if (!is.vector(color.node) || length(color.node) != 
                2) 
                stop("'color.node' must be a vector of length 2.", 
                  call. = FALSE)
        }
        else {
            stop("'color.node' must be a vector of length 2.", 
                call. = FALSE)
        }
    }
    if (any(!sapply(color.node, function(x) {
        tryCatch(is.matrix(col2rgb(x)), error = function(e) FALSE)
    }))) 
        stop("'color.node' must be a character vector of recognized colors.", 
            call. = FALSE)
    if (any(class.object %in% object.blocks)) {
        if (is.null(shape.node)) 
            shape.node = "circle"
        if (is.vector(shape.node)) {
            if (length(shape.node) == 1) 
                shape.node = rep(shape.node, length(blocks))
        }
        if (!is.list(shape.node)) {
            if (!is.vector(shape.node) || length(shape.node) != 
                length(blocks)) 
                stop("'shape.node' must be a character vector of length ", 
                  length(blocks), ".", call. = FALSE)
        }
        else {
            stop("'shape.node' must be a numeric vector of length ", 
                length(blocks), ".", call. = FALSE)
        }
        if (!all(shape.node %in% c("none", "circle", "rectangle"))) 
            stop("elements of 'shape.node' must be from {'none', 'circle', 'rectangle'}.", 
                call. = FALSE)
        if (is.null(names(shape.node))) 
            names(shape.node) = blocks
    }
    else {
        if (is.null(shape.node)) 
            shape.node = c("circle", "rectangle")
        if (length(shape.node) == 1) {
            if (shape.node == "none") {
                shape.node = c("none", "none")
            }
        }
        if (!is.list(shape.node)) {
            if (!is.vector(shape.node) || length(shape.node) != 
                2) 
                stop("'shape.node' must be a vector of length 2.", 
                  call. = FALSE)
        }
        else {
            stop("'shape.node' must be a vector of length 2.", 
                call. = FALSE)
        }
        if (!all(shape.node %in% c("none", "circle", "rectangle"))) 
            stop("elements of 'shape.node' must be from {'none', 'circle', 'rectangle'}.", 
                call. = FALSE)
    }
    if (!is.null(cex.node.name)) {
        if (!is.finite(cex.node.name) || cex.node.name < 0 || 
            length(cex.node.name) > 1) {
            stop("'cex.node.name' must be a non-negative numerical value.", 
                call. = FALSE)
        }
    }
    if (length(color.edge) < 2 && (!is(color.edge, "function"))) 
        stop("'color.edge' must be a vector of length larger than or equal to 2.", 
            call. = FALSE)
    if ((length(color.edge)%%2) != 0 && (!is(color.edge, "function")) && 
        isTRUE(symkey)) 
        stop("'color.edge' must be a vector of length an even number if 'symkey = TRUE'.", 
            call. = FALSE)
    if (any(!sapply(color.edge, function(x) {
        tryCatch(is.matrix(col2rgb(x)), error = function(e) FALSE)
    }))) 
        stop("'color.edge' must be a character vector of recognized colors.", 
            call. = FALSE)
    if (length(lty.edge) > 2) 
        stop("\"lty.edge\" must a character vector of up to 2 entries from\n        'solid', 'dashed', 'dotted', 'dotdash', 'longdash', twodash' or 'blank'.\n        see ?network", 
            call. = FALSE)
    if (length(lty.edge) == 1) 
        lty.edge = c(lty.edge, lty.edge)
    choices = c("solid", "dashed", "dotted", "dotdash", "longdash", 
        "twodash", "blank")
    lty.edge = choices[pmatch(lty.edge, choices, duplicates.ok = TRUE)]
    if (any(is.na(lty.edge))) 
        stop("'lty.edge' should be from 'solid', 'dashed', 'dotted', 'dotdash', 'longdash', twodash' or 'blank'.", 
            call. = FALSE)
    if (length(lwd.edge) > 2) 
        stop("'lwd.edge' must be a vector of up to 2 positive numbers.\n        See ?network")
    if (length(lwd.edge) == 1) 
        lwd.edge = c(lwd.edge, lwd.edge)
    if (length(lwd.edge) != 2 || any(!is.finite(lwd.edge)) || 
        any(lwd.edge <= 0)) 
        stop("'lwd.edge' must be positive.")
    if (!is.logical(show.edge.labels)) 
        stop("'show.edge.labels' must be a logical constant (TRUE or FALSE).", 
            call. = FALSE)
    if (!is.finite(cex.edge.label) || cex.edge.label < 0 || length(cex.edge.label) > 
        1) 
        stop("'cex.edge.label' must be a non-negative numerical value.", 
            call. = FALSE)
    if (!is.logical(show.color.key)) 
        stop("'show.color.key' must be a logical constant (TRUE or FALSE).", 
            call. = FALSE)
    if (!is.logical(symkey)) 
        stop("'symkey' must be a logical constant (TRUE or FALSE).", 
            call. = FALSE)
    if (length(keysize) != 2 || any(!is.finite(keysize))) 
        stop("'keysize' must be a numeric vector of length 2.", 
            call. = FALSE)
    if (length(keysize.label) != 1 || any(!is.finite(keysize))) 
        stop("'keysize' must be a numeric vector of length 1.", 
            call. = FALSE)
    if (!is.logical(interactive)) 
        stop("'interactive' must be a logical constant (TRUE or FALSE).", 
            call. = FALSE)
    if (!is.null(layout.fun) && !is(layout.fun, "function")) 
        stop("'layout.fun' must be a valid layout function.", 
            call. = FALSE)
    if (!(any(class.object %in% object.blocks))) 
        w = as.vector(t(mat))
    if (round(max(abs(w)), 2) == 0) 
        stop("There is no correlation between these blocks whith these components.", 
            "Try a different value of 'comp'.", call. = FALSE)
    if (!is.numeric(cutoff) || cutoff < 0 || cutoff > 1) 
        stop("'cutoff' should be a numeric between 0 and 1", 
            call. = FALSE)
    if (cutoff > max(abs(w))) 
        stop("You have chosen a high cutoff value of ", cutoff, 
            " which is greaer than the max value in the similarity matrix which is ", 
            round(max(abs(w)), 2), call. = FALSE)
    if (any(class.object %in% object.blocks)) {
        group = NULL
        temp = lapply(mat$X, function(x) colnames(x))
        for (i in 1:length(temp)) {
            group = c(group, rep(names(temp)[i], length(temp[[i]])))
        }
        nodes = data.frame(name = unlist(temp), group = group)
    }
    else if (any(class.object %in% object.pls)) {
        w = as.vector(t(mat))
        Xn = sum(keep.X)
        Yn = sum(keep.Y)
        node.X = row.names
        node.Y = col.names
        row.names.plot = row.names.plot[keep.X]
        col.names.plot = col.names.plot[keep.Y]
        nodes = data.frame(name = c(node.X, node.Y), group = c(rep("x", 
            Xn), rep("y", Yn)))
        node.X = rep(node.X, each = Yn)
        node.Y = rep(node.Y, Xn)
    }
    else {
        node.X = row.names
        node.Y = col.names
        nodes = data.frame(name = c(node.X, node.Y), group = c(rep("x", 
            p), rep("y", q)))
        node.X = rep(node.X, each = q)
        node.Y = rep(node.Y, p)
    }
    relations = data.frame(from = node.X, to = node.Y, weight = w)
    id = bin.color(w, cutoff = cutoff, breaks = breaks, col = color.edge, 
        symkey = symkey)
    col.id = id$bin
    color.edge = id$col[col.id]
    idx = (abs(w) >= cutoff)
    relations = relations[idx, ]
    color.edge = color.edge[idx]
    gR = graph.data.frame(relations, directed = FALSE, vertices = nodes)
    V(gR)$label.color = "black"
    V(gR)$label.family = "sans"
    if (any(class.object %in% object.blocks)) {
        V(gR)$label = unlist(block.var.names)
        j = 1
        for (i in blocks) {
            V(gR)$color[V(gR)$group == i] = color.node[j]
            V(gR)$shape[V(gR)$group == i] = shape.node[j]
            if (shape.node[j] == "none") {
                V(gR)$label.color[V(gR)$group == i] = paste0(substr(color.node[j], 
                  1, 7), "FF")
            }
            j = j + 1
        }
    }
    else {
        V(gR)$label = c(row.names.plot, col.names.plot)
        V(gR)$color = color.node[1]
        V(gR)$color[V(gR)$group == "y"] = color.node[2]
        V(gR)$shape = shape.node[1]
        if (shape.node[1] == "none") {
            V(gR)$label.color = paste0(substr(color.node[1], 
                1, 7), "FF")
        }
        V(gR)$shape[V(gR)$group == "y"] = shape.node[2]
        if (shape.node[2] == "none") {
            V(gR)$label.color[V(gR)$group == "y"] = paste0(substr(color.node[2], 
                1, 7), "FF")
        }
    }
    if (show.edge.labels) 
        E(gR)$label = round(E(gR)$weight, 2)
    E(gR)$label.color = "black"
    E(gR)$color = color.edge
    E(gR)$lty = lty.edge[1]
    E(gR)$lty[E(gR)$weight < 0] = lty.edge[2]
    E(gR)$width = lwd.edge[1]
    E(gR)$width[E(gR)$weight < 0] = lwd.edge[2]
    gR = delete.vertices(gR, which(degree(gR) == 0))
    lwid = c(keysize[1], 4)
    lhei = c(keysize[2], 4)
    lmat = matrix(c(1, 2, 3, 4), 2, 2, byrow = TRUE)
    nc = length(id$col)
    x = seq(0, 1, length = nc + 2)
    z.mat = seq(0, 1, length = nc + 1)
    z.mat = matrix(z.mat, ncol = 1)
    if ((id$lim[1] < -cutoff) & (id$lim[2] < cutoff)) {
        xv = c(0, x[nc + 1])
        lv = round(c(id$lim[1], -cutoff), 2)
        col = c(id$col, "white")
    }
    if ((id$lim[1] > -cutoff) & (id$lim[2] > cutoff)) {
        xv = c(x[2], 1)
        lv = round(c(cutoff, id$lim[2]), 2)
        col = c("white", id$col)
    }
    if ((id$lim[1] < -cutoff) & (id$lim[2] > cutoff)) {
        idn = max(which(id$breaks < 0))
        idp = min(which(id$breaks > 0))
        xv = c(0, x[idn + 1], x[idp], 1)
        lv = round(c(id$lim[1], -cutoff, cutoff, id$lim[2]), 
            2)
        col = c(id$col[1:idn], "white", id$col[(idn + 1):nc])
    }
    nn = vcount(gR)
    V(gR)$label.cex = ifelse(is.null(cex.node.name), 1, cex.node.name)
    E(gR)$label.cex = min(2.25 * cex.edge.label/log(nn), 1)
    cex0 = 2 * V(gR)$label.cex
    def.par = par(no.readonly = TRUE)
    dev.new()
    par(pty = "s", mar = c(0, 0, 0, 0), mfrow = c(1, 1))
    upr.bnd <- 190 * (graph.scale^2) + 10
    plot(1:upr.bnd, 1:upr.bnd, type = "n", axes = FALSE, xlab = "", 
        ylab = "")
    cha = V(gR)$label
    cha = paste("", cha, "")
    xh = strwidth(cha, cex = ifelse(size.node == 0.5, cex0, size.node * 
        4)) * 1.5 * (2 - graph.scale)
    yh = strheight(cha, cex = ifelse(size.node == 0.5, cex0, 
        size.node * 4)) * 3 * (2 - graph.scale)
    V(gR)$size = xh
    V(gR)$size2 = yh
    dev.off()
    if (is.null(layout.fun)) {
        l = layout.fruchterman.reingold(gR, weights = (2 - abs(E(gR)$weight))) # 1 -> 2
    }
    else {
        l = layout.fun(gR)
    }
    if (isTRUE(!interactive)) {
        if (isTRUE(show.color.key)) {
            l = layout.fruchterman.reingold(gR, weights = (2 - # 1 -> 2
                abs(E(gR)$weight)))
        }
        else {
            l = layout.fun(gR)
        }
        if (isTRUE(!interactive)) {
            if (isTRUE(show.color.key)) {
                layout(lmat, widths = lwid, heights = lhei, respect = FALSE)
                par(mar = c(5, 4, 2, 1), cex = 0.75)
                image(z.mat, col = col, xaxt = "n", yaxt = "n")
                box()
                par(usr = c(0, 1, 0, 1))
                axis(1, at = xv, labels = lv, cex.axis = keysize.label)
                title("Color key", font.main = 1, cex.main = keysize.label)
                par(def.par)
                par(new = TRUE)
            }
            par(pty = "s", mar = c(0, 0, 0, 0), mfrow = c(1, 
                1))
            plot(gR, layout = l)
            par(def.par)
        }
    }
    gE.none = FALSE
    if (isTRUE(interactive)) {
        min.cut = cutoff
        max.cut = max(abs(w))
        cutoff.old = cutoff
        dev.new(width = 5, height = 2.7, xpos = -1)
        def.par = par(no.readonly = TRUE)
        cuts = seq(0, 1, length = 21)
        par(mai = c(0.25, 0.15, 0.3, 0.15), bg = gray(0.95))
        layout(matrix(c(0, 1, 0), ncol = 1, nrow = 3), widths = 1, 
            heights = c(0.25, 1, 0.25))
        plot(cuts, type = "n", rep(0, 21), xlab = "", ylab = "", 
            xlim = c(-0.1, 1.1), axes = FALSE)
        title("cutoff control", cex.main = 1.9, font.main = 1)
        text(0.5, -0.6, "value", cex = 1.5)
        text(0, -0.6, round(min.cut, 2), cex = 1.4)
        text(1, -0.6, round(max.cut, 2), cex = 1.4)
        mtext(min.cut, side = 1, line = -1, outer = FALSE, cex = 0.95)
        rect(-0.1, -0.3, -0.02, 0.3, col = "white")
        rect(1.02, -0.3, 1.1, 0.3, col = "white")
        points(1.06, 0, pch = 3, cex = 2.4)
        lines(c(-0.085, -0.035), c(0, 0))
        for (i in seq(0, 1, length = 21)) lines(c(i, i), c(-0.22, 
            0.2))
        x = pos = 0
        rect(-0.01, -0.045, x, 0.04, col = "red")
        rect(x, -0.045, 1.01, 0.04, col = "white")
        bar.dev = dev.cur()
        dev.new()
        net.dev = dev.cur()
        if (isTRUE(show.color.key)) {
            layout(lmat, widths = lwid, heights = lhei, respect = FALSE)
            par(mar = c(5, 4, 2, 1), cex = 0.75)
            image(z.mat, col = col, xaxt = "n", yaxt = "n")
            box()
            par(usr = c(0, 1, 0, 1))
            axis(1, at = xv, labels = lv, cex.axis = keysize.label)
            title("Color key", font.main = 1, cex.main = keysize.label)
            par(def.par)
            par(new = TRUE)
        }
        par(pty = "s", mar = c(0, 0, 0, 0))
        plot(gR, layout = l)
        par(def.par)
        old.pos = -1
        repeat {
            dev.set(bar.dev)
            z = locator(1, type = "n")
            x = z[[1]]
            y = z[[2]]
            if (is.null(z)) 
                break
            if (0 <= x & x <= 1 & -0.22 <= y & y <= 0.22) {
                rect(0, -0.045, x, 0.04, col = "red")
                rect(x, -0.045, 1.01, 0.04, col = "white")
                pos = x
            }
            if (1.02 <= x & x <= 1.1 & -0.3 <= y & y <= 0.3) {
                x = pos + 0.05
                idx = which.min(abs(cuts - x))
                x = cuts[idx]
                pos = x
                rect(0, -0.045, x, 0.04, col = "red")
                rect(x, -0.045, 1.01, 0.04, col = "white")
            }
            if (-0.1 <= x & x <= -0.02 & -0.3 <= y & y <= 0.3) {
                x = pos - 0.05
                idx = which.min(abs(cuts - x))
                x = cuts[idx]
                pos = x
                rect(0, -0.045, x, 0.04, col = "red")
                rect(x, -0.045, 1.01, 0.04, col = "white")
            }
            if (old.pos != pos) {
                old.pos = pos
                rect(0.4, -0.8, 0.6, -1.5, col = gray(0.95), 
                  border = NA)
                cutoff = (max.cut - min.cut) * pos + min.cut
                mtext(round(cutoff, 3), side = 1, line = -1, 
                  cex = 0.9)
                dev.set(net.dev)
                if (cutoff >= cutoff.old) {
                  supp.edge = E(gR)[abs(E(gR)$weight) < cutoff]
                  gE = delete.edges(gR, supp.edge)
                  gE = delete.vertices(gE, which(degree(gE) == 
                    0))
                  nn = vcount(gE)
                  V(gR)$label.cex = min(2.5 * cex.node.name/log(nn), 
                    1)
                  E(gR)$label.cex = min(2.25 * cex.edge.label/log(nn), 
                    1)
                  cex0 = 2 * V(gE)$label.cex
                  def.par = par(no.readonly = TRUE)
                  par(pty = "s", mar = c(0, 0, 0, 0))
                  plot(1:100, 1:100, type = "n", xaxt = "n")
                  cha = V(gE)$label
                  cha = paste("", cha, "")
                  xh = strwidth(cha, cex = cex0) * 1.5
                  yh = strheight(cha, cex = cex0) * 3
                  V(gE)$size = xh
                  V(gE)$size2 = yh
                  par(def.par)
                  if (is.null(layout.fun)) {
                    l = layout.fruchterman.reingold(gE, weights = (2 - # 1 -> 2
                      abs(E(gE)$weight)))
                  }
                  else {
                    l = layout.fun(gE)
                  }
                  if (isTRUE(show.color.key)) {
                    layout(lmat, widths = lwid, heights = lhei, 
                      respect = FALSE)
                    par(mar = c(5, 4, 2, 1), cex = 0.75)
                    image(z.mat, col = col, xaxt = "n", yaxt = "n")
                    box()
                    par(usr = c(0, 1, 0, 1))
                    axis(1, at = xv, labels = lv, cex.axis = keysize.label)
                    title("Color key", font.main = 1, cex.main = keysize.label)
                    par(def.par)
                    par(new = TRUE)
                  }
                  par(pty = "s", mar = c(0, 0, 0, 0))
                  plot(gE, layout = l)
                  par(def.par)
                  cutoff.old = cutoff
                }
                else {
                  supp.edge = E(gR)[abs(E(gR)$weight) < cutoff]
                  gE = delete.edges(gR, supp.edge)
                  gE = delete.vertices(gE, which(degree(gE) == 
                    0))
                  nn = vcount(gE)
                  V(gR)$label.cex = min(2.5 * cex.node.name/log(nn), 
                    1)
                  E(gR)$label.cex = min(2.25 * cex.edge.label/log(nn), 
                    1)
                  cex0 = 2 * V(gE)$label.cex
                  def.par = par(no.readonly = TRUE)
                  par(pty = "s", mar = c(0, 0, 0, 0))
                  plot(1:100, 1:100, type = "n", xaxt = "n")
                  cha = V(gE)$label
                  cha = paste("", cha, "")
                  xh = strwidth(cha, cex = cex0) * 1.5
                  yh = strheight(cha, cex = cex0) * 3
                  V(gE)$size = xh
                  V(gE)$size2 = yh
                  par(def.par)
                  if (is.null(layout.fun)) {
                    l = layout.fruchterman.reingold(gE, weights = (2 - # 1 -> 2
                      abs(E(gE)$weight)))
                  }
                  else {
                    l = layout.fun(gE)
                  }
                  if (isTRUE(show.color.key)) {
                    layout(lmat, widths = lwid, heights = lhei, 
                      respect = FALSE)
                    par(mar = c(5, 4, 2, 1), cex = 0.75)
                    image(z.mat, col = col, xaxt = "n", yaxt = "n")
                    box()
                    par(usr = c(0, 1, 0, 1))
                    axis(1, at = xv, labels = lv, cex.axis = keysize.label)
                    title("Color key", font.main = 1, cex.main = keysize.label)
                    par(def.par)
                    par(new = TRUE)
                  }
                  par(pty = "s", mar = c(0, 0, 0, 0))
                  plot(gE, layout = l)
                  par(def.par)
                  cutoff.old = cutoff
                }
                gE.none = TRUE
            }
        }
        if (gE.none != FALSE) 
            gR = gE
    }
    res = list(gR = gR)
    if (any(class.object %in% object.blocks)) {
        l = 1
        for (i in 1:(length(blocks) - 1)) {
            for (j in (i + 1):length(blocks)) {
                M_block[[l]][abs(M_block[[l]]) < cutoff] = 0
                res[paste("M", blocks[i], blocks[j], sep = "_")] = list(M_block[[l]])
                l = l + 1
            }
        }
    }
    else {
        mat[abs(mat) < cutoff] = 0
        res$M = mat
    }
    res$cutoff = cutoff
    if (!is.null(save)) 
        dev.off()
    return(invisible(res))
}


```



```{r}

# BiocManager::install("mixOmicsTeam/mixOmics@devel")
library(mixOmics)


# if (!requireNamespace("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")

# BiocManager::install("BiocParallel")
library(BiocParallel)



```

# Data: Enz, Met, rna, pro and Pph

```{r}

fp = file.path('..', 'input')


fnfp = file.path('..', 'input')


fn = 'Enzymomics_only_24_shortnames_used_diablo.csv'
E = read.delim(file = file.path(fp, fn), 
               header = TRUE, 
               sep = ",", 
               quote = NULL,
               dec = ".", 
               fill = TRUE, 
               comment.char = '@')

fn = 'Metabolomics_only_24_shortnames_diablo.csv'
M = read.delim(file = file.path(fp, fn), 
               header = TRUE, 
               sep = ",", 
               quote = NULL,
               dec = ".", 
               fill = TRUE, 
               comment.char = '@')

fn = 'phosphoProteomics_filtered_sig_24_shortnames_diablo.csv'
PP = read.delim(file = file.path(fp, fn), 
               header = TRUE, 
               sep = ",", 
               quote = NULL,
               dec = ".", 
               fill = TRUE, 
               comment.char = '@')

fn = 'Proteomics_filtered_sig_24_with short names.csv'
P = read.delim(file = file.path(fp, fn), 
               header = TRUE, 
               sep = ",", 
               quote = NULL,
               dec = ".", 
               fill = TRUE, 
               comment.char = '@')

fn = 'rnaseq_diablo_filtered_shortnames.csv'
R = read.delim(file = file.path(fp, fn), 
               header = TRUE, 
               sep = ",", 
               quote = NULL,
               dec = ".", 
               fill = TRUE, 
               comment.char = '@')

all(P$treatment == E$treatment)
all(P$treatment == M$treatment)
all(P$treatment == PP$treatment)
all(P$treatment == R$treatment)




```


## log2


```{r, 00_densities}

range(P[, -1], na.rm = TRUE)
range(M[, -1], na.rm = TRUE)
range(E[, -1], na.rm = TRUE)
range(PP[, -1], na.rm = TRUE)
range(R[, -1], na.rm = TRUE)

# range(log(M[, -1], 10), na.rm = TRUE)
# range(log(E[, -1], 10), na.rm = TRUE)

range(log(M[, -1], 2), na.rm = TRUE)
range(log(E[, -1], 2), na.rm = TRUE)


# PBSmodelling::plotDens(M[, -1])
# PBSmodelling::plotDens(log(M[, -1], 2))


tmp = tidyr::gather(M, var, measurement, colnames(M)[2]:colnames(M)[ncol(M)], 
                    factor_key = FALSE)

range(tmp$measurement, na.rm = TRUE)

ggplot(tmp, aes(x=measurement, fill=treatment)) +
  geom_density(alpha = 0.1)

ggplot(tmp, aes(x=measurement, fill=var)) +
  geom_density(alpha = 0.1) +
  theme(legend.position="none")


tmp$measurement = log(tmp$measurement, 2)

range(tmp$measurement, na.rm = TRUE)

ggplot(tmp, aes(x=measurement, fill=treatment)) +
  geom_density(alpha = 0.1)

ggplot(tmp, aes(x=measurement, fill=var)) +
  geom_density(alpha = 0.1) +
  theme(legend.position="none")



tmp = tidyr::gather(E, var, measurement, colnames(E)[2]:colnames(E)[ncol(E)], 
                    factor_key = FALSE)
range(tmp$measurement, na.rm = TRUE)

ggplot(tmp, aes(x=measurement, fill=treatment)) +
  geom_density(alpha = 0.1)

ggplot(tmp, aes(x=measurement, fill=var)) +
  geom_density(alpha = 0.1) +
  theme(legend.position="none")


tmp$measurement = log(tmp$measurement, 2)
range(tmp$measurement, na.rm = TRUE)

ggplot(tmp, aes(x=measurement, fill=treatment)) +
  geom_density(alpha = 0.1)

ggplot(tmp, aes(x=measurement, fill=var)) +
  geom_density(alpha = 0.1) +
  theme(legend.position="none")



```



```{r}


M[, -1] = log(M[, -1], 2)
E[, -1] = log(E[, -1], 2)

```

see also <https://mixomics.org/methods/missing-values/>

<https://mixomics.org/faq/common-issues-and-errors/>

A large degree of sparsity within your data, such that there is a large number of missing or zero values. One suggestion would be to remove all features that have more than a certain threshold (eg. 20%) of zero/missing value. If this does not work, refer to the Missing Values page on how to impute these values manually. 



```{r}

E.imp = impute.nipals(X = E[, -1], ncomp = 10)


```


numeric table

```{r}

data = list(Enz = as.matrix(E[, -1]),
            Met = as.matrix(M[, -1]),
            Pph = as.matrix(P[, -1]),
            Pro = as.matrix(PP[, -1]),
            Rna = as.matrix(R[, -1]))





```

<https://mixomics.org/mixdiablo/diablo-tcga-case-study/>

```{r}


lapply(data, dim) # check their dimensions
Y = factor(E$treatment, levels = unique(E$treatment))
summary(Y)


```

# Pairwise PLS Comparisons

Circle Correlation Plots for pairwise PLS models 


Only displays the top 24 features for each dimension, subsetting by those with a correlation above 0.5

```{r, 01_paired-plotVar}

# 24 is the smalles data set here
no = min(unlist(lapply(data, ncol)))
list.keepX = c(no, no) # select arbitrary values of features to keep
list.keepY = c(no, no)

pls1 <- spls(data[["Enz"]], data[["Met"]], 
             keepX = list.keepX, keepY = list.keepY) 
pls2 <- spls(data[["Enz"]], data[["Pph"]], 
             keepX = list.keepX, keepY = list.keepY)
pls3 <- spls(data[["Met"]], data[["Pph"]], 
             keepX = list.keepX, keepY = list.keepY)
pls4 <- spls(data[["Enz"]], data[["Pro"]], 
             keepX = list.keepX, keepY = list.keepY)
pls5 <- spls(data[["Enz"]], data[["Rna"]], 
             keepX = list.keepX, keepY = list.keepY)
pls6 <- spls(data[["Met"]], data[["Pro"]], 
             keepX = list.keepX, keepY = list.keepY)
pls7 <- spls(data[["Met"]], data[["Rna"]], 
             keepX = list.keepX, keepY = list.keepY)
pls8 <- spls(data[["Pph"]], data[["Pro"]], 
             keepX = list.keepX, keepY = list.keepY)
pls9 <- spls(data[["Pph"]], data[["Rna"]], 
             keepX = list.keepX, keepY = list.keepY)
pls10 <- spls(data[["Pro"]], data[["Rna"]], 
             keepX = list.keepX, keepY = list.keepY)
# plot features of first PLS
plotVar(pls1, cutoff = 0.5, title = "(a) enzymomics vs metabolomics", 
        legend = c("enzymomics", "metabolomics"), 
        var.names = TRUE, style = 'graphics', 
        pch = c(16, 17), cex = c(0.5,0.5), 
        col = c('darkorchid', 'lightgreen'))
plotVar(pls1, cutoff = 0.5, title = "(a) enzymomics vs metabolomics", 
        legend = c("enzymomics", "metabolomics"), 
        var.names = FALSE, style = 'graphics', 
        pch = c(16, 17), cex = c(2,2), 
        col = c('darkorchid', 'lightgreen'))

# plot features of second PLS
plotVar(pls2, cutoff = 0.5, title = "(b) enzymomics vs phosphoproteomics", 
        legend = c("enzymomics", "phosphoproteomics"), 
        var.names = TRUE, style = 'graphics', 
        pch = c(16, 17), cex = c(0.5,0.5), 
        col = c('darkorchid', 'blue'))
plotVar(pls2, cutoff = 0.5, title = "(b) enzymomics vs phosphoproteomics", 
        legend = c("enzymomics", "phosphoproteomics"), 
        var.names = FALSE, style = 'graphics', 
        pch = c(16, 17), cex = c(2,2), 
        col = c('darkorchid', 'blue'))


# plot features of third PLS
plotVar(pls3, cutoff = 0.5, title = "(c) metabolomics vs phosphoproteomics", 
        legend = c("metabolomics", "phosphoproteomics"), 
        var.names = TRUE, style = 'graphics', 
        pch = c(16, 17), cex = c(0.5,0.5), 
        col = c('lightgreen', 'blue'))
plotVar(pls3, cutoff = 0.5, title = "(c) metabolomics vs phosphoproteomics", 
        legend = c("metabolomics", "phosphoproteomics"), 
        var.names = FALSE, style = 'graphics', 
        pch = c(16, 17), cex = c(2,2), 
        col = c('lightgreen', 'blue'))

plotVar(pls4, cutoff = 0.5, title = "(a) enzymomics vs proteomicss", 
        legend = c("enzymomics", "proteomics"), 
        var.names = TRUE, style = 'graphics', 
        pch = c(16, 17), cex = c(0.5,0.5), 
        col = c('darkorchid', 'red'))
plotVar(pls4, cutoff = 0.5, title = "(a) enzymomics vs proteomics", 
        legend = c("enzymomics", "proteomics"), 
        var.names = FALSE, style = 'graphics', 
        pch = c(16, 17), cex = c(2,2), 
        col = c('darkorchid', 'red'))

plotVar(pls5, cutoff = 0.5, title = "(a) enzymomics vs RNA-seq", 
        legend = c("enzymomics", "RNA-seq"), 
        var.names = TRUE, style = 'graphics', 
        pch = c(16, 17), cex = c(0.5,0.5), 
        col = c('darkorchid', 'orange'))
plotVar(pls5, cutoff = 0.5, title = "(a) enzymomics vs RNA-seq", 
        legend = c("enzymomics", "RNA-seq"), 
        var.names = FALSE, style = 'graphics', 
        pch = c(16, 17), cex = c(2,2), 
        col = c('darkorchid', 'orange'))

plotVar(pls6, cutoff = 0.5, title = "(c) metabolomics vs proteomics", 
        legend = c("metabolomics", "proteomics"), 
        var.names = TRUE, style = 'graphics', 
        pch = c(16, 17), cex = c(0.5,0.5), 
        col = c('lightgreen', 'red'))
plotVar(pls6, cutoff = 0.5, title = "(c) metabolomics vs proteomics", 
        legend = c("metabolomics", "proteomics"), 
        var.names = FALSE, style = 'graphics', 
        pch = c(16, 17), cex = c(2,2), 
        col = c('lightgreen', 'red'))

plotVar(pls7, cutoff = 0.5, title = "(c) metabolomics vs RNA-seq", 
        legend = c("metabolomics", "RNA-seq"), 
        var.names = TRUE, style = 'graphics', 
        pch = c(16, 17), cex = c(0.5,0.5), 
        col = c('lightgreen', 'orange'))
plotVar(pls7, cutoff = 0.5, title = "(c) metabolomics vs RNA-seq", 
        legend = c("metabolomics", "RNA-seq"), 
        var.names = FALSE, style = 'graphics', 
        pch = c(16, 17), cex = c(2,2), 
        col = c('lightgreen', 'orange'))

plotVar(pls8, cutoff = 0.5, title = "(c) phosphoproteomics vs proteomics", 
        legend = c("phosphoproteomics", "proteomics"), 
        var.names = TRUE, style = 'graphics', 
        pch = c(16, 17), cex = c(0.5,0.5), 
        col = c('blue', 'red'))
plotVar(pls8, cutoff = 0.5, title = "(c) phosphoproteomics vs phosphoproteomics", 
        legend = c("phosphoproteomics", "proteomics"), 
        var.names = FALSE, style = 'graphics', 
        pch = c(16, 17), cex = c(2,2), 
        col = c('blue', 'red'))

plotVar(pls9, cutoff = 0.5, title = "(c) phosphoproteomics vs RNA-seq", 
        legend = c("phosphoproteomics", "RNA-seq"), 
        var.names = TRUE, style = 'graphics', 
        pch = c(16, 17), cex = c(0.5,0.5), 
        col = c('blue', 'orange'))
plotVar(pls9, cutoff = 0.5, title = "(c) phosphoproteomics vs RNA-seq", 
        legend = c("phosphoproteomics", "RNA-seq"), 
        var.names = FALSE, style = 'graphics', 
        pch = c(16, 17), cex = c(2,2), 
        col = c('blue', 'orange'))

plotVar(pls10, cutoff = 0.5, title = "(c) proteomics vs RNA-seq", 
        legend = c("proteomics", "RNA-seq"), 
        var.names = TRUE, style = 'graphics', 
        pch = c(16, 17), cex = c(0.5,0.5), 
        col = c('red', 'orange'))
plotVar(pls10, cutoff = 0.5, title = "(c) proteomics vs RNA-seq", 
        legend = c("proteomics", "RNA-seq"), 
        var.names = FALSE, style = 'graphics', 
        pch = c(16, 17), cex = c(2,2), 
        col = c('red', 'orange'))


```

## correlation

```{r}

c
cor(pls1$variates$X, pls1$variates$Y) # calculate correlation of Enz and Met
cor(pls2$variates$X, pls2$variates$Y) # calculate correlation of Enz and Pph
cor(pls3$variates$X, pls3$variates$Y) # calculate correlation of Met and Pph
cor(pls4$variates$X, pls4$variates$Y)
cor(pls5$variates$X, pls5$variates$Y)
cor(pls6$variates$X, pls6$variates$Y)
cor(pls7$variates$X, pls7$variates$Y)
cor(pls8$variates$X, pls8$variates$Y)
cor(pls9$variates$X, pls9$variates$Y)
cor(pls10$variates$X, pls10$variates$Y)

```


# Design



```{r}

design = matrix(0.5, ncol = length(data), nrow = length(data), # for square matrix filled with 0.1s
                dimnames = list(names(data), names(data)))
diag(design) = 0 # set diagonal to 0s

design

```

#  the initial DIABLO 


```{r}

basic.diablo.model = block.splsda(X = data, Y = Y, ncomp = 5, design = design) # form basic DIABLO model

```


# Tuning the number of components

Choosing the number of components in block.plsda using perf() with 10  10-fold CV function 


10-fold Cross-validation & leave-one-out cross-validation: <https://stats.stackexchange.com/questions/154830/10-fold-cross-validation-vs-leave-one-out-cross-validation>

```{r, 02_no-components}
## ---- fig.cap = "FIGURE 2: Choosing the number of components in `block.plsda` using `perf()` with 10  10-fold CV function in the `breast.TCGA` study. Classification error rates (overall and balanced, see Section 7.3) are represented on the y-axis with respect to the number of components on the x-axis for each prediction distance presented in PLS-DA"----

perf.diablo = perf(basic.diablo.model, 
                   dist = 'all', 
                   cpus = 1,
                   validation = 'loo', 
                   progressBar = TRUE,
                   folds = 10, 
                   nrepeat = 10) # run component number tuning with repeated CV

plot(perf.diablo) # plot output of tuning

```


##  set the optimal ncomp value


```{r}


ncomp = perf.diablo$choice.ncomp$WeightedVote["Overall.BER", "centroids.dist"] # set the optimal ncomp value
perf.diablo$choice.ncomp$WeightedVote # show the optimal choice for ncomp for each dist metric


```

# Tuning the number of features

We choose the optimal number of variables to select in each data set using the tune.block.splsda() function, for a grid of keepX values for each type of omics. Note that the function has been set to favour a relatively small signature while allowing us to obtain a sufficient number of variables for downstream validation and/or interpretation.


## setup cluster - use SnowParam() on Widnows

<https://rdrr.io/bioc/BiocParallel/man/BiocParallelParam-class.html>


bplapply erro solution <https://support.bioconductor.org/p/133353/#133356>

```{r}

options(SnowParam=SnowParam(workers=1))
BPPARAM  = bpparam()
BPPARAM <- BiocParallel::SnowParam(workers = parallel::detectCores()-1)




```


Crate tune.MIR with tune.block.splsda

and save RData (beacuse it takes cca 1h)

then comment these lines

and just load RData

10-fold Cross-validation & leave-one-out cross-validation: <https://stats.stackexchange.com/questions/154830/10-fold-cross-validation-vs-leave-one-out-cross-validation>

```{r}

# set grid of values for each component to test


 # test.keepX = list (Met = c(5:9, seq(10, 15, 2), seq(20,24,4)),
 #                    Enz = c(5:9, seq(10, 15, 2), seq(20,24,4)),
 #                    Pph = c(5:9, seq(10, 15, 2), seq(20,24,4)))

x <- list()
for (i in 1:length(data)){
x[[i]] <- c( seq(5,min(24, ncol(data[[i]])) ,2))
}
names(x) <- names(data)
test.keepX  <-  x
test.keepX



fp = file.path('..', 'output')
fn = '01_DIABLO-Enz-Met-Pph-Pro-Rna_tune.MIR.RData'


# You have provided a sequence of keepX of length: 10 for block Met and 10 for block Enz and 10 for block Pph.
# This results in 1000 models being fitted for each component and each nrepeat, this may take some time to run, be patient!
# Because of a too high number of 'folds' required, 2 folds were randomly assigned no data: the number of 'folds' is reduced to 8
# .....
 
  t1 = proc.time()
 # 
 #run the feature selection tuning
 library(stats)
 tune.MIR = tune.block.splsda(X = data, Y = Y,
                              ncomp = ncomp,
                              test.keepX = test.keepX,
                              design = design,
                              validation = 'loo',
                             folds = 10, nrepeat = 1, # Leave-One-Out validation does not need to be repeated
                              BPPARAM  = BPPARAM ,
                              dist = "centroids.dist",
                              max.iter = 200)
 t2 = proc.time()
 running_time = t2 - t1; running_time
 save(tune.MIR, file = file.path(fp, fn))
 
load(file.path(fp, fn))


list.keepX = tune.MIR$choice.keepX
list.keepX



```


The number of features to select on each component is returned in


```{r}



list.keepX = tune.MIR$choice.keepX # set the optimal values of features to retain
list.keepX


```

# Final model

```{r}


# set the optimised DIABLO model
final.diablo.model = block.splsda(X = data, Y = Y, ncomp = ncomp
                          , keepX = list.keepX
                          , design = design)

## Design matrix has changed to include Y; each block will be
##             linked to Y.



```



```{r}

# the features selected from components
for (comp in 1:ncomp){
cat("\nComponent ", comp,":\n")
for(i in 1:length(data)){
cat(names(data)[i],"\n")
print(selectVar(final.diablo.model, comp = comp)[[i]]$name)
}
}



```


# Sample plots


```{r, 03_Sample-Component-plots}


for(comp in 1:ncomp){
plotDiablo(final.diablo.model, ncomp = comp)
title(paste("Component",comp), adj=0.1, line=-1, outer=TRUE)
}

```


```{r, 04_Sample-plots}


plind <- plotIndiv(final.diablo.model, ind.names = FALSE, legend = TRUE,
          title = 'DIABLO Sample Plots'
          , ellipse = TRUE,
          comp = c(1,3)
          )

plind <- plotIndiv(final.diablo.model, ind.names = FALSE, legend = TRUE,
          title = 'DIABLO Sample Plots'
          , ellipse = TRUE,
          comp = c(1,2)
          )
plind <- plotIndiv(final.diablo.model, ind.names = FALSE, legend = TRUE,
          title = 'DIABLO Sample Plots'
          , ellipse = TRUE,
          comp = c(2,3)
          )


```


```{r, 05_Arrow-sample-plot}

plotArrow(final.diablo.model, ind.names = FALSE, legend = TRUE,
          title = paste(groups,collapse=", ")
          )



```


# Variable plots

```{r, 06_Correlation-circle-plots}

if(length(data)==3) pick <- 1:3 else pick <- c(4,1:3)
cols <- c('orange1', 'brown1', 'lightgreen',"lightblue")[pick]
pchs <- c(16, 17, 15, 18)[pick]
plotVar(final.diablo.model, var.names = FALSE,
        style = 'graphics', legend = TRUE
        , pch = pchs, cex = rep(2,length(data))
        , col = cols
)

plotVar(final.diablo.model, var.names = TRUE,
        style = 'graphics', legend = TRUE
        , pch = pchs, cex = rep(0.5,length(data))
        , col = cols
)

```


```{r, 07_circosPlot_0.50}

cutoff = 0.50

circosPlot(final.diablo.model, cutoff = cutoff, line = TRUE,
           color.blocks= cols,
           color.cor = c(3,2), size.labels = 1
           , xpd=TRUE)




```


```{r, 07_circosPlot_0.90}

cutoff = 0.90

circosPlot(final.diablo.model, cutoff = cutoff, line = TRUE,
           color.blocks= cols,
           color.cor = c(3,2), size.labels = 1
           , xpd=TRUE)




```



```{r, 07_circosPlot_0.75}

cutoff = 0.75

circosPlot(final.diablo.model, cutoff = cutoff, line = TRUE,
           color.blocks= cols,
           color.cor = c(3,2), size.labels = 1
           , xpd=TRUE)




```




# Relevance networks





# more plots

issue <https://github.com/mixOmicsTeam/mixOmics/issues/45>

```{r, 08_plot-Loadings}


for(i in 1:ncomp)
plotLoadings(final.diablo.model, comp = i, contrib = 'max', method = 'median')


# plotLoadings encountered margin errors. Ensure feature names are not too long (see 'name.var' argument) and the 'Plots' pane is cleared and enlargened.


traceback()


```

```{r, Clustered-Image-Maps}

cimfn <- "cim.png"
png(cimfn, res = 600, width = 8000, height = 8000)
cimDiablo(final.diablo.model, size.legend=0.7, margins = c(15, 15))
dev.off()



```


```{r, 09_Clustered-Image-Maps}
#heatmaps
cimDiablo(final.diablo.model, margins = c(13, 13), size.legend=0.8)

```


```{r, 10_network}
#cutoff for correlations

network(final.diablo.model, 
        blocks = c(1,2,3),
        color.node = c('darkorchid', 'brown1', 'lightgreen'), 
        cutoff = 0.50,
        size.node = 0.05)

network(final.diablo.model, 
        blocks = c(1,2,3),
        color.node = c('darkorchid', 'brown1', 'lightgreen'), 
        cutoff = 0.90,
        size.node = 0.10)


network(final.diablo.model, 
        blocks = c(1,2,3),
        color.node = c('darkorchid', 'brown1', 'lightgreen'), 
        cutoff = 0.75,
        size.node = 0.10)

```


```{r, 11_ROC}

par(mfrow=c(2,2))
for(i in 1:length(data))
auc.splsda = auroc(final.diablo.model, roc.block = names(data[i]),
                   roc.comp = 2, print = FALSE)



```




no cut-off

```{r}

net = network(final.diablo.model, 
              blocks = seq(1, ncomp, 1),
              color.node = c('darkorchid', 'brown1', 'lightgreen'), 
              cutoff = 0,
              size.node = 0.05,
              plot.graph = FALSE,
              layout.fun = igraph::layout_in_circle)

typeof(net)
mode(net)
names(net)
summary(net)

net$cutoff
# net$gR
range(net$M_Enz_Met)
range(net$M_Enz_Pph)
range(net$M_Met_Pph)

names(dimnames(net$M_Enz_Met)) = c('from','to')
a = as.data.frame(
  as.table(net$M_Enz_Met),
  responseName = 'CC'
  )
a$fromType = 'Enz'
a$toType = 'Met'
a$ID1 = gsub('_Enz$', '', a$from)
a$ID2 = gsub('_Met$', '', a$to)

names(dimnames(net$M_Enz_Pph)) = c('from','to')
b = as.data.frame(
  as.table(net$M_Enz_Pph),
  responseName = 'CC'
  )
b$fromType = 'Enz'
b$toType = 'Pph'
b$ID1 = gsub('_Enz$', '', b$from)
b$ID2 = gsub('_Pph$', '', b$to)

names(dimnames(net$M_Met_Pph)) = c('from','to')
e = as.data.frame(
  as.table(net$M_Met_Pph),
  responseName = 'CC'
  )
e$fromType = 'Met'
e$toType = 'Pph'
e$ID1 = gsub('_Met$', '', e$from)
e$ID2 = gsub('_Pph$', '', e$to)

merged = rbind(a, b, e)
dim(merged)
summary(net$gR)
gl = igraph::as_edgelist(net$gR, names = TRUE)
dim(gl)

```


<!-- # Final analysis -->

<!-- c/p from ```GitHub/UJI-Omics-phd/_I_UJI-Omics-phd/_S_multiOmicsIntegra/_A_10_mixOmics-R/scripts/03_Step5.html``` -->


# per groups

## NM only

```{r}

ind = grep('NM', E[, 1])
data.NM = purrr::map(data, function(x) {
  x[ind,]
  })



res.NM <- block.splsda(X = data.NM
                       , Y = as.factor(E[ind, 1])
                       , ncomp = ncomp
                       , keepX = list.keepX
                       , design = design
                       )

range(res.NM$weights)

net.NM = network(res.NM, 
                 blocks = seq(1, ncomp, 1),
                 color.node = c('darkorchid', 'brown1', 'lightgreen'), 
                 cutoff = 0,
                 size.node = 0.05,
                 plot.graph = FALSE,
                 layout.fun = igraph::layout_in_circle)

names(dimnames(net.NM$M_Enz_Met)) = c('from','to')
a = as.data.frame(
  as.table(net.NM$M_Enz_Met),
  responseName = 'CC'
  )
a$fromType = 'Enz'
a$toType = 'Met'
a$ID1 = gsub('_Enz$', '', a$from)
a$ID2 = gsub('_Met$', '', a$to)

names(dimnames(net.NM$M_Enz_Pph)) = c('from','to')
b = as.data.frame(
  as.table(net.NM$M_Enz_Pph),
  responseName = 'CC'
  )
b$fromType = 'Enz'
b$toType = 'Pph'
b$ID1 = gsub('_Enz$', '', b$from)
b$ID2 = gsub('_Pph$', '', b$to)

names(dimnames(net.NM$M_Met_Pph)) = c('from','to')
e = as.data.frame(
  as.table(net.NM$M_Met_Pph),
  responseName = 'CC'
  )
e$fromType = 'Met'
e$toType = 'Pph'
e$ID1 = gsub('_Met$', '', e$from)
e$ID2 = gsub('_Pph$', '', e$to)

merged.NM = rbind(a, b, e)


```


## AM only

```{r}


ind = grep('AM', E[, 1])
data.AM = purrr::map(data, function(x) {
  x[ind,]
  })

res.AM <- block.splsda(X = data.AM
                       , Y = as.factor(E[ind, 1])
                       , ncomp = ncomp
                       , keepX = list.keepX
                       , design = design
                       )

range(res.AM$weights)

net.AM = network(res.AM, 
                blocks = seq(1, ncomp, 1),
                 color.node = c('darkorchid', 'brown1', 'lightgreen'), 
                 cutoff = 0,
                 size.node = 0.05,
                 plot.graph = FALSE,
                 layout.fun = igraph::layout_in_circle)

names(dimnames(net.AM$M_Enz_Met)) = c('from','to')
a = as.data.frame(
  as.table(net.AM$M_Enz_Met),
  responseName = 'CC'
  )
a$fromType = 'Enz'
a$toType = 'Met'
a$ID1 = gsub('_Enz$', '', a$from)
a$ID2 = gsub('_Met$', '', a$to)

names(dimnames(net.AM$M_Enz_Pph)) = c('from','to')
b = as.data.frame(
  as.table(net.AM$M_Enz_Pph),
  responseName = 'CC'
  )
b$fromType = 'Enz'
b$toType = 'Pph'
b$ID1 = gsub('_Enz$', '', b$from)
b$ID2 = gsub('_Pph$', '', b$to)

names(dimnames(net.AM$M_Met_Pph)) = c('from','to')
e = as.data.frame(
  as.table(net.AM$M_Met_Pph),
  responseName = 'CC'
  )
e$fromType = 'Met'
e$toType = 'Pph'
e$ID1 = gsub('_Met$', '', e$from)
e$ID2 = gsub('_Pph$', '', e$to)

merged.AM = rbind(a, b, e)


```

## inf only

```{r}


ind = grep('inf', E[, 1])
data.inf = purrr::map(data, function(x) {
  x[ind,]
  })

res.inf <- block.splsda(X = data.inf
                       , Y = as.factor(E[ind, 1])
                       , ncomp = ncomp
                       , keepX = list.keepX
                       , design = design
                       )

range(res.inf$weights)

net.inf = network(res.inf, 
                blocks = seq(1, ncomp, 1),
                 color.node = c('darkorchid', 'brown1', 'lightgreen'), 
                 cutoff = 0,
                 size.node = 0.05,
                 plot.graph = FALSE,
                 layout.fun = igraph::layout_in_circle)

names(dimnames(net.inf$M_Enz_Met)) = c('from','to')
a = as.data.frame(
  as.table(net.inf$M_Enz_Met),
  responseName = 'CC'
  )
a$fromType = 'Enz'
a$toType = 'Met'
a$ID1 = gsub('_Enz$', '', a$from)
a$ID2 = gsub('_Met$', '', a$to)

names(dimnames(net.inf$M_Enz_Pph)) = c('from','to')
b = as.data.frame(
  as.table(net.inf$M_Enz_Pph),
  responseName = 'CC'
  )
b$fromType = 'Enz'
b$toType = 'Pph'
b$ID1 = gsub('_Enz$', '', b$from)
b$ID2 = gsub('_Pph$', '', b$to)

names(dimnames(net.inf$M_Met_Pph)) = c('from','to')
e = as.data.frame(
  as.table(net.inf$M_Met_Pph),
  responseName = 'CC'
  )
e$fromType = 'Met'
e$toType = 'Pph'
e$ID1 = gsub('_Met$', '', e$from)
e$ID2 = gsub('_Pph$', '', e$to)

merged.inf = rbind(a, b, e)


```

## uninf only

```{r}


ind = grep('inf', E[, 1], invert = TRUE)
data.uninf = purrr::map(data, function(x) {
  x[ind,]
  })

res.uninf <- block.splsda(X = data.uninf
                       , Y = as.factor(E[ind, 1])
                       , ncomp = ncomp
                       , keepX = list.keepX
                       , design = design
                       )

range(res.uninf$weights)

net.uninf = network(res.uninf, 
                blocks = seq(1, ncomp, 1),
                 color.node = c('darkorchid', 'brown1', 'lightgreen'), 
                 cutoff = 0.000000,
                 size.node = 0.05,
                 plot.graph = FALSE,
                 layout.fun = igraph::layout_in_circle)

names(dimnames(net.uninf$M_Enz_Met)) = c('from','to')
a = as.data.frame(
  as.table(net.uninf$M_Enz_Met),
  responseName = 'CC'
  )
a$fromType = 'Enz'
a$toType = 'Met'
a$ID1 = gsub('_Enz$', '', a$from)
a$ID2 = gsub('_Met$', '', a$to)

names(dimnames(net.uninf$M_Enz_Pph)) = c('from','to')
b = as.data.frame(
  as.table(net.uninf$M_Enz_Pph),
  responseName = 'CC'
  )
b$fromType = 'Enz'
b$toType = 'Pph'
b$ID1 = gsub('_Enz$', '', b$from)
b$ID2 = gsub('_Pph$', '', b$to)

names(dimnames(net.uninf$M_Met_Pph)) = c('from','to')
e = as.data.frame(
  as.table(net.uninf$M_Met_Pph),
  responseName = 'CC'
  )
e$fromType = 'Met'
e$toType = 'Pph'
e$ID1 = gsub('_Met$', '', e$from)
e$ID2 = gsub('_Pph$', '', e$to)

merged.uninf = rbind(a, b, e)


```



```{r}

colnames(merged)[3] = paste(colnames(merged)[3], 'all', sep = '_')
colnames(merged.AM)[3] = paste(colnames(merged.AM)[3], 'AM', sep = '_')
colnames(merged.NM)[3] = paste(colnames(merged.NM)[3], 'NM', sep = '_')
colnames(merged.inf)[3] = paste(colnames(merged.inf)[3], 'inf', sep = '_')
colnames(merged.uninf)[3] = paste(colnames(merged.uninf)[3], 'uninf', sep = '_')

dim(merged)
dim(merged.AM)
dim(merged.NM)
dim(merged.inf)
dim(merged.uninf)

toWrite = NULL
toWrite = merge(merged, 
                merged.AM, 
                #merged.NM,
                #merged.inf,
                #merged.uninf,
                by = c('from', 'to', "fromType", "toType", "ID1", "ID2"),
                all.x = TRUE, all.y = TRUE)
dim(toWrite)
toWrite = merge(toWrite,
                merged.NM,
                #merged.inf,
                #merged.uninf,
                by = c('from', 'to', "fromType", "toType", "ID1", "ID2"),
                all.x = TRUE, all.y = TRUE)
dim(toWrite)
toWrite = merge(toWrite,
                merged.inf,
                #merged.uninf,
                by = c('from', 'to', "fromType", "toType", "ID1", "ID2"),
                all.x = TRUE, all.y = TRUE)
dim(toWrite)
toWrite = merge(toWrite,
                merged.uninf,
                by = c('from', 'to', "fromType", "toType", "ID1", "ID2"),
                all.x = TRUE, all.y = TRUE)
dim(toWrite)
table(duplicated(toWrite[, 1:2]))

intersect(grep('IDH', toWrite$from), grep('sophoroside', toWrite$to))
intersect(grep('IDH', merged$to), grep('sophoroside', merged$from))
intersect(grep('IDH', gl[, 2]), grep('sophoroside', gl[, 1]))
intersect(grep('IDH', gl[, 1]), grep('sophoroside', gl[, 2]))


table(toWrite$from %in% toWrite$to)
table(toWrite$fromType %in% toWrite$toType)
rs = rowSums(is.na(toWrite[, 7:ncol(toWrite)]))
table(rs)
toWrite[which(rs == 4), ]


```


write cor

```{r}

fp = file.path('..', 'output')
fn = '01_DIABLO-Enz_Met_Pph-both.txt'

write.table(merged, 
            file = file.path(fp, fn), 
            append = FALSE, 
            quote = FALSE, 
            sep = "\t",
            eol = "\n", 
            na = "NA", 
            dec = ".", 
            row.names = FALSE,
            col.names = TRUE, 
            qmethod = "escape",
            fileEncoding = "UTF-8")


```





# Session

```{r}

# devtools::session_info()
sessionInfo()

```

