---
title: "03_geneID_to_CKN-Ath"
author: "zagor"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_caption: yes
    self_contained: yes
    fig_width: 16
    fig_height: 9
    toc: true
    toc_float:
      toc_collapsed: false
    toc_depth: 5
    number_sections: yes
    theme: flatly
    highlight: tango
  pdf_document:
    toc: yes
    toc_depth: '2'
  word_document:
    toc: yes
    toc_depth: '2'
editor_options:
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(#dev = c('pdf', 'png'),  # this embeds pdf and crates scrolable blocks
                      dev = c('png'), 
                      fig.align = 'center', 
                      fig.height = 9, 
                      fig.width = 16 ,
                      warning = FALSE, message = FALSE
                      )
# options(knitr.table.format = "html")

```


```{r,  echo=TRUE, warning=FALSE, message=FALSE}

rm(list = ls(all = TRUE))
gc()

set.seed(123456)


`%nin%` = Negate(`%in%`)

library(magrittr)


```


# p-value cut-off

```{r}

pval.cutoff = 0.05


```

# input

## transation table

```{r}

fpt = file.path('..', 'input')
fnt = 'tomato-arabidopsis-translation_dummy.txt'


```

## limma results

```{r}


fpp = file.path('..', '..', '_A_02_DEtranscriptomics-R', 'output', 'limma-out')
fl = list.files(fpp)


```


## PSS nodes (Ath based)

```{r}

fs = 'old_CKN_nodes.txt'

```

## read transation table

```{r}


NGS.ID.to.PIS = read.table(file.path(fpt, fnt),
                           header = TRUE, 
                           sep = "\t", 
                           quote = NULL,
                           dec = ".", 
                           stringsAsFactors = FALSE,
                           na.strings = "NA",
                           fill = TRUE,
                           comment.char = "#")

NGS.ID.to.PIS = NGS.ID.to.PIS[!duplicated(NGS.ID.to.PIS), ]
# tomato
colnames(NGS.ID.to.PIS)[1] = 'GeneID'
# arabidopsis is he key
NGS.ID.to.PIS = data.table::data.table(NGS.ID.to.PIS, key = "AthID") 
dim(NGS.ID.to.PIS)

```


## read IDs that are present in PSS

```{r}


sel = read.table(file.path(fpt, fs),
                 header = TRUE, 
                 sep = "\t", 
                 quote = NULL,
                 dec = ".", 
                 stringsAsFactors = FALSE,
                 na.strings = "NA",
                 fill = TRUE,
                 comment.char = "#")

table(duplicated(sel$geneID))

# shrink table to run faster
# NGS.ID.to.PIS = NGS.ID.to.PIS[NGS.ID.to.PIS$AthID %in% sel$geneID, ]
table(duplicated(NGS.ID.to.PIS$AthID))

```


## merge limma output and filter by adjusted p-value

take care of column names that u grep

```{r}

organism.NGS.files = NGS.ID.to.PIS

for (i in fl) {
  
  tmp = read.table(file = file.path(fpp, i),
                   header = TRUE, 
                   sep = "\t", 
                   quote = NULL,
                   dec = ".", 
                   stringsAsFactors = FALSE,
                   na.strings = "NA",
                   fill = TRUE,
                   comment.char = "#")  
  
  ind = c(grep("GeneID", colnames(tmp)), grep("adj.P.Val", colnames(tmp)), grep("logFC", colnames(tmp)))
  tmp = tmp[,ind]
  ind = which(tmp[,2] < pval.cutoff)
  tmp = tmp[ind,c(1,3)]
  colnames(tmp)[2] = i
  # tmp[, 1] = gsub('s', 'S', tmp[, 1])
  organism.NGS.files = merge(organism.NGS.files, tmp, by = "GeneID", all.x = TRUE, sort = FALSE)
}


dim(organism.NGS.files)






```


# prioritisation

## selec first column where logFC filtered values start

```{r}

logFC.starts.from.col = 3

```


## prioritise - most responsive tomatoID to AthID

```{r}

# check that numerical columns are cumeric
# lapply(organism.NGS.files, class)

# unique Ath nodes in PSS
clades = sort(unique(organism.NGS.files$AthID))
tmp = NULL

for (i in clades) {
  ind = which(organism.NGS.files$AthID == i)
  if (length(ind) > 1) { # (length(ind > 1)) for found, not length(ind) > 1 for multiple (1 or more, doesnt matter here)
    mysubset = organism.NGS.files[ind, ]
    e = sapply(1:nrow(mysubset), function(x) sum(!is.na(mysubset[x, logFC.starts.from.col:ncol(mysubset)])))
    f1 = sapply(1:nrow(mysubset), 
            function(x) ifelse(all(is.na(mysubset[x, logFC.starts.from.col:ncol(mysubset)])), 
                                0, 
                                mean(colMeans(abs(mysubset[x, logFC.starts.from.col:ncol(mysubset)]), na.rm = TRUE), na.rm = TRUE)))
    f2 = sapply(1:nrow(mysubset), 
            function(x) ifelse(all(is.na(mysubset[x, logFC.starts.from.col:ncol(mysubset)])), 
                                0, 
                                max(abs(mysubset[x, logFC.starts.from.col:ncol(mysubset)]), na.rm = TRUE)))
    g = which(e == max(e))
    h1 = which(f1[g] == max(f1[g]))
    h2 = which(f2[g] == max(f2[g]))
    k2 = ifelse(length(h1) > 1, h2, h1)
    k = ifelse(length(g) > 1, g[k2], g)
    
    tmp = rbind(tmp, mysubset[k,])
  } else { # no match
    tmp = rbind(tmp, organism.NGS.files[ind,]) # take first
  }
}

nrow(tmp) == length(clades)
clades.organism.NGS =  tmp

# we can have the same tomato matching different Ath
table(duplicated(clades.organism.NGS$GeneID))
# but we should have AtID multiple times
table(duplicated(clades.organism.NGS$AthID))




```



```{r}

openxlsx::write.xlsx(clades.organism.NGS, 
                     file.path("..", "output", "03b_tomato-ath_prioritised.xlsx"), 
                     asTable = TRUE, overwrite = TRUE)


```


# write output in limma style

```{r}

fpo = file.path('..', 'output', 'Ath-DiNAR-input')
dir.create(fpo)
fpo = file.path('..', 'output', 'Ath-DiNAR-input', 'prioritised')
dir.create(fpo)
getGenes = clades.organism.NGS$GeneID
length(getGenes)
# remove ones that have only 0 (or NA) per all conditions
j = !sapply(1:nrow(clades.organism.NGS), 
            function(x) all(is.na(clades.organism.NGS[x, logFC.starts.from.col:ncol(clades.organism.NGS)])))
getGenes = getGenes[j]
length(getGenes)


for (i in fl) {
  
  tmp = read.table(file = file.path(fpp, i),
                   header = TRUE, 
                   sep = "\t", 
                   quote = NULL,
                   dec = ".", 
                   stringsAsFactors = FALSE,
                   na.strings = "NA",
                   fill = TRUE,
                   comment.char = "#")
  
  # Solyc vs solyc
  # tmp$GeneID = gsub('s', 'S', tmp$GeneID)
  ind = match(getGenes, tmp$GeneID)
  tmp = tmp[ind, ]
  table(duplicated(clades.organism.NGS$AthID))
  temp = merge(tmp, clades.organism.NGS[, 1:2], by = "GeneID", all.x = FALSE, all.y = FALSE,  sort = FALSE)
  # be sure not to have duplicated IDs
  temp = temp[!duplicated(temp), ]
  temp = dplyr::arrange(temp, GeneID, AthID)
  table(duplicated(temp$AthID))
  
  write.table(x = temp, file = file.path(fpo, i), 
              append = FALSE, 
              quote = FALSE, 
              sep = "\t",
              eol = "\n", 
              na = "NA", 
              dec = ".", 
              row.names = FALSE,
              col.names = TRUE)
}


```


# sessionInfo

```{r}

sessionInfo()

```

